Extracting Angular Observables without a Likelihood\\and Applications to Rare Decays

### 2015-01-22 draft.tex
Changes after another discussion with Fred

\begin{abstract}
  \fred{Our goal is to obtain a complete set} of angular observables
  arising in a generic multi-body process without the need to carry
  out a likelihood fit of the angular distribution to the measured
  events. Instead, we \fred{apply the method of moments} that relies
  both on the orthogonality of angular functions and the estimation of
  integrals by Monte Carlo techniques. \fred{The main advantage of
    this method is that the joint distribution of all observables
    given the data can be easily extracted, even for very few
    events. Our main contribution is to show explicitly how systematic
    uncertainties from detector-resolution and acceptance effects can
    be incorporated in a general algorithm. Finally, we present the
    necessary process-dependent formulae needed for direct
    implementation of the method for several rare decays currently of
    interest.}
\end{abstract}

@@

The initial motivation for studying what we wish to call the
\emph{method of moments} is the determination of angular observables
in the rare decay FCNC-mediated decay $\bar{B}\to \bar{K}^*(\to
\bar{K}\pi)\ell^+\ell^-$. However, we emphasize that the method we
describe in this letter is general, and it applies to arbitrary decay
or scattering processes. We find a previous work \cite{Dighe:1998vk}
advocates this method chiefly for the determination of angular
observables in non-leptonic $B$ decays, but also mentions the
applicability to semileptonic decays. We improve upon this previous
work by studying correlations \fred{really? because of higher $l$}
between the angular observables, and by showing explicitly how to
incorporate uncertainties due to acceptance effects. Moreover, we show
that the method of moments has two major advantages over the usual
approach based on likelihood fits:
\begin{enumerate}
    \item Likelihood fits have convergence problems for a small number of
        events, and can require reparametrizations and/or approximations
        for a successful fit to the signal PDF, see e.g. \cite{Egede:2008uy}.
        The method of moments has no such convergence problems.
    \item The results of likelihood fits suffer from systematic uncertainties
        and systematic bias (see e.g. \cite{Egede:2008uy}), especially in but not limited
        to cases in which the PDFs do not faithfully model the underlying physics. We will
        show that the method of moments does not suffer a very particular type of
        mismodelling that arises from introducing a cutoff in a partial wave expansion.
    \item Using the method of moments, the joint probability distribution of the
        angular observables converges towards a multivariate Gaussian distribution.
        This allows an easy transfer of correlation information from the experiments to
        interested theorists.
\end{enumerate}

We continue with basic definitions that pertain to angular observables, and our
results in the subsequent sections. Let $\thvec = (\theta_1, \dots \theta_N)$ denote the set of all
angles, and let $\nuvec = (\nu_1, \dots \nu_M)$ denote the set of all
other non-angular kinematic variables needed to fully specify the
final state of the process under study. For example, $\nuvec$ may
include invariant masses or center-of-mass energies. We define an
angular observable $S_i$ as a coefficient in the probability density
function (PDF) of the process by means of
\begin{align}
    \label{eq:def-P}
    P(\nuvec, \thvec) \equiv \sum_i S_i(\nuvec) \times f_i(\thvec)\,.
\end{align}
Here, the dependence on the $n$ decay angles $\thvec$ has been
explicitly factored out in terms of the angular functions
$\{f_i(\thvec)\}$. We assume there exists a dual basis of functions
$\{\dual{f}_i(\thvec)\}$ such that the orthonormality relations
\begin{equation}
    \label{eq:def-ortho-rel}
    \int_\Omega \rmdx{^N \theta} \dual{f}_i(\thvec) f_j(\thvec)  = \delta_{ij}
\end{equation}
hold with $\Omega$ representing the full angular phase space relevant
to the process.  For particle decays, $P$ is generally expressed in
terms of the fully differential decay width,
\begin{align}
    \label{eq:def-P-decay}
    P(\nuvec, \thvec) \equiv \frac{1}{\Gamma}\frac{\rmdx{^{N+M}\Gamma}}{\rmdx{\nu_1} \dots \rmdx{\nu_M} \rmdx{\theta_1} \dots \rmdx{\theta_N}}\,,
\end{align}
where $\Gamma$ is the total decay width. For a scattering process, one can similarly use
\begin{align}
    \label{eq:def-P-scattering}
    P(\nuvec, \thvec) \equiv \frac{1}{\sigma}\frac{\rmdx{^{N+M}\sigma}}{\rmdx{\nu_1} \dots \rmdx{\nu_M} \rmdx{\theta_1} \dots \rmdx{\theta_N}}\,,
\end{align}
where the total cross section $\sigma$ is used for the
normalization. Since the determination of the total decay width or
total cross section can be quite difficult, we emphasize that
different normalizations for $P$ can be used.  For instance, the total
decay width (or cross section) of the process of interest can be
replaced by the corresponding quantity of a control-channel
process. This change of normalization is equivalent to a linear rescaling of
the angular observables $\lbrace S_i\rbrace$; thus ratios or similar suitable combinations
of the angular observables are invariant.\\

% Our method is closely related to Pearson's method of moments
% \citeneeded{}. One draws samples from a PDF $P(x | \nuvec)$ that
% depends on parameters $\nuvec$. Then one approximates the population
% (central) moments $x^k$ by the sample moments
Our method is an extension of the classical method of moments with
orthogonal functions~\cite[sec. 8.2]{James:2006zz}. The only difference
is that conventionally the angular functions are assumed
\emph{self-dual}, $\dual{f}_i = f_i$. However, it suffices that the
system of angular functions $\{f_i(\thvec)\}$ \emph{can} be transformed into
an orthonormal basis. It turns out to be much more convenient to work
in the basis of Legendre polynomials but the statistical properties
are equivalent for both approaches provided that one replace $f_i \to
\dual{f}_i$ appropriately. Using the ansatz
\begin{equation}
  \label{eq:dual-ansatz}
  \dual{f}_i = \sum_{j} a_{ij} f_j \,,
\end{equation}
the dual basis needs to be worked out case by case through solving the
linear system of equations (\ref{eq:def-ortho-rel}). For a selection
of hadron decays with a $b$ quark in the initial state and two leptons
in the final state, we list the dual bases in a series of appendices
\ref{app:btokll} through \ref{app:lambdabtolambdall}. Note that a
similar analysis was done in~\cite{Dighe:1998vk} for the decays $B \to
J/\psi \phi$ and $B \to J/\psi K^{*}$.

@@

The orthonormality relations \refeq{def-ortho-rel} imply that a single angular observable $S_i$
can be projected out of the full PDF $P$ by means of
\begin{equation}
    \label{eq:det-Pi-analytical}
    S_i(\nuvec) = \int_{\Omega} \rmdx{^N \theta}  P(\nuvec, \thvec) \dual{f}_i(\thvec)\,.
\end{equation}
where $\lbrace \dual{f}_i \rbrace$ denotes a \emph{dual basis} of angular functions. In general, $\lbrace \dual{f}_i \rbrace$ may differ from $\lbrace f_i \rbrace$. This is the
case for our selection of applications in appendices \ref{app:btokll} through \ref{app:lambdabtolambdall}.\\

@@

Integration over the non-angular variables yields
\begin{equation}
    \langle S_i\rangle
    \equiv \int \rmdx{^M \nu} S_i(\nuvec)
    = \int \rmdx{^M \nu} \left[\int_{\Omega} \rmdx{^N \theta} P(\nuvec,\thvec) \dual{f}_i(\thvec) \right].
\end{equation}
The remainder of this section describes the method of moments, in
which we replace the analytical integration by Monte Carlo (MC)
estimates.  The central tenet of MC integration is the fact that the
expectation value $E_P[g]$ of some function $g(x)$ under the
probability density $P(x)$,
\begin{equation}
    E_P[g] \equiv \int \rmdx{x} P(x) g(x),
\end{equation}
can be approximated by the consistent and unbiased
estimator $\est{E_P[g]}$~\cite[sec. 8.2]{James:2006zz}
\begin{equation}
    \label{eq:mc-id}
    E_P[g] \to \widehat{E_P[g]} \equiv \frac{1}{K} \sum_{k=1}^K g(x^{(k)}) \,,\,    x^{(k)} \sim P
\end{equation}
due to the strong law of large numbers for $K \to \infty$, assuming
that the variates $x^{(k)}$, $k = 1, \dots, K$, are distributed as
$P$.
Throughout this letter we denote all MC estimators with a wide hat.\\

Application of \refeq{mc-id} then yields
\begin{equation}
    \langle S_i\rangle \to \widehat{\langle S_i\rangle} = \frac{1}{K} \sum_{k=1}^{K} \dual{f}_i(x^{(k)})\,.
\end{equation}
It is often of interest to obtain observables integrated over certain
bins of $\nuvec$. We define
\begin{align}
    \langle S_i\rangle_{\vec{a},\vec{b}}
    & \equiv \int_{\vec{a}}^{\vec{b}} \rmdx{^M \nu} S_i(\nuvec)\\
    & = \int_{\vec{a}}^{\vec{b}} \rmdx{^M \nu} \left[\int_{\Omega} \rmdx{^N\theta} P(\nuvec,\thvec) \dual{f}_i(\thvec)  \right]\\
    & = \int \rmdx{^M \nu} \left[\int_{\Omega} \rmdx{^N\theta}  P(\nuvec,\thvec) \dual{f}_i(\thvec)
        \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,
        \right],
\end{align}
where the argument of the indicator function $ \mathbf{1}(\vec{a} \le
\nuvec \le \vec{b})$ is to be interpreted componentwise.
Application of \refeq{mc-id} immediately yields
\begin{equation}
    \label{eq:bin-importance}
    \widehat{\langle S_i\rangle_{\vec{a},\vec{b}}}
    = \frac{1}{K} \sum_{k=1}^{K} \dual{f}_i(x^{(k)})         \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,.
\end{equation}
For notational simplicity, let us forget about the $\nuvec$
integration and consider only $S_i$. In the limit $K \to \infty$, the
central limit theorem (CLT) implies that the random vector
\begin{equation}
  \label{eq:angular-obs-vec}
  \vecest{S} \equiv (S_1, \dots, S_i\,,
  \dots)
\end{equation}
follows a multivariate Gaussian distribution $\mathcal{N}( \vec{S},
\widehat{\Sigma})$ centered on the true value $\vec{S}$ with the
covariance $\widehat{\Sigma}_{ij}$ estimated as
\begin{equation}
    \cov[S_i,S_j] \to \est{\cov}[{S}_i, {S}_j]
        = \frac{1}{K - 1} \sum_{k=1}^{K} \big[\dual{f}_i(x^{(k)}) - \est{ S_i}\big]\,\big[\dual{f}_j(x^{(k)}) - \widehat{ S_j}\big]\,.
\end{equation}
In our physics applications, the parameter space is compact and each
$\dual{f}_i$ is bounded. Hence the requisites for the most basic
version of the CLT to hold --- finite mean and covariance of
$\dual{f}_i$ --- are automatically satisfied.  We would like to
emphasize that the sample covariance should \fred{``Should'' as in it
  does in practice? Reference toy studies below} rapidly converge
towards the true covariance matrix.  We propose to declare convergence
\fred{I'd prefer: check that skewness and kurtosis are sufficiently
  small} once estimators for skewness and kurtosis are sufficiently
small,

\begin{equation}
    \est{\operatorname{Skew}}[S_i,S_j] \leq B_S\,\,\text{ and }\,\,\est{\kurt}[S_i,S_j] \leq B_K\,.
\end{equation}
where the limits $B_S$ and $B_K$ should be determined from process-specific dedicated toy studies.\\

Compared to the usual maximum-likelihood approach, we find for the method of moments:
\begin{enumerate}
  \item The angular observable ${S_i}$ can be determined
  independently of any other observable ${S_j}$. It is therefore
  much more robust to physics assumptions needed to define the full
  likelihood. In particular, this means one does not have to be
  specific regarding the form of new-physics contributions; in fact,
  one does not even need to be able to explicitly formulate the
  likelihood at all.
\item It is superior for a small number of samples $N$. Likelihood
  fits tend to be numerically unstable if lots of parameters need to
  be estimated from sparse data. This is more severe if the mode of
  the likelihood is near the boundary of the physically allowed
  region.\fred{Is there some LHCb note we could cite?} In many of the
  decays of interest, there are only say dozens\todo{check!} of events recorded.
\item The estimate is unbiased for any $N$. In contrast, the
  maximum-likelihood estimate has a bias of order
  $1/N$~\cite{Cox:1968}. In practice, one should keep in mind the
  bias-variance trade-off: it is a well known phenomenon that removing
  the bias usually leads to an increase in variance of the sampling
  distribution of the estimator~\cite[sec. 7.3]{James:2006zz}. From a
  Bayesian decision-theory point of view, both contribute similarly to
  the expected loss associated with deciding on just one value of the
  unknown parameter. It is therefore highly questionable to reduce
  only the bias~\cite[sections 13.8,17.2]{jaynes:2003}. In fact, for
  the results discussed below in \refsec{numerics}, the likelihood
  fits --- if they converge --- produce uncertainties (10--15)\%
  smaller than those from the method of moments.
  \item The approximate multivariate Gaussian distribution of $\vecest{S}$
  allows easier and more correct transfer of the information in the
  data to interested theorists for more accurate fits of
  standard-model and new-physics parameters \cite{Altmannshofer:2013foa,
    Descotes-Genon:2013wba,Bobeth:2013soa},
  or for more precise predictions of optimized observables; e.g.,
  \cite{Egede:2008uy,Egede:2010zc,Bobeth:2010wg,Becirevic:2011bp,
    Bobeth:2012vn,Matias:2012xw,DescotesGenon:2012zf} for definitions
  of such optimized observables in $B\to K^*\ell^+\ell^-$ decays,
  \cite{Faller:2013dwa} for application to the decay $B\to
  \pi\pi\ell^-\bar\nu_\ell$, and \cite{Boeer:2014xx} for observables
  in $\Lambda_b\to\Lambda(\to N\pi)\ell^+\ell^-$). While the
  likelihood also approaches a multivariate Gaussian as $K \to
  \infty$, the two methods differ in their utility as input for
  theorists if $\widehat{\vec{S}}$ is not well
  inside the physical region. For example, suppose there are two
  angular observables that are constrained to a triangular region by
  phase-space or unitarity arguments as \fred{Could add a sketch. It
    is supposed to resemble $A_{\text{FB}},F_L$ for $B \to K^* \ell
    \ell$.}
  \begin{equation}
    \label{eq:constr-ex}
    |S_1| + S_2 \le 1, \, S_1 \in [-1,1],\, S_2 \in [0,1] \,.
  \end{equation}
  It may (and often does) happen in practice that $\widehat{\vec{S}}$
  is close or even outside the allowed region such that a significant
  part of the probability mass covers unphysical values. In a Bayesian
  fit, one would take $\mathcal{N}(\vecest{S} | \vec{S},
  \widehat{\Sigma})$ as the sampling distribution of the ``data'', and
  simply set a uniform prior on the triangle in the $\vec{S}$ plane
  defined by \refeq{constr-ex} to have a well defined problem. This
  could be trivially combined with other independent information in a
  global fit. Someone with a different physics model might have to
  consider a
  different physical region, and could incorporate it just as easily.\\

  For a likelihood fit, the constraint needs to be part of the
  analysis performed by the experimental collaboration, and the
  resulting likelihood as a function of $\vec{S}$ is distinctly not
  Gaussian.  For experimentalists, communicating such a result has
  proved to be challenging due to technical reasons (data formats,
  size etc.) and the fear that theorists misinterprete the many
  nuisance parameters (from control channels, detector calibrations
  etc.) that are also part of the likelihood. This leads to the
  undesirable situation that that only the mode and standard errors
  are reported, and theorists often include the results as independent
  measurements with a Gaussian distribution and disregard the boundary
  problem as well as correlations altogether.

@@

In several interesting processes we might only have an approximate
result for $P$.  In this section, we focus on one particular class of
mismodelling of the signal PDF: the angular-momentum cutoff in partial
wave expansions. This mismodelling potentially affects a large number
of decays and scattering processes. For the sake of clarity we take
the interesting class of four-body decays $B\to P_1 P_2 \ell_1
\bar\ell_2$ \fred{How general is this? The plot implies that the
  dimeson and the dilepton do not have a common vertex. But this
  requires an intermediate state, e.g., $K^{*}$}as an
example\footnote{This includes the rare $b\to s$ mediated $B$ decay $B
  \to K\pi\ell^+\ell^-$, and the $V_{ub}$ suppressed decay $B\to
  \pi\pi\ell^+\bar\nu_\ell$. For both examples the PDF $P$ is known in
  the small-width approximation and when assuming a pure $P$-wave,
  resonant final state. An extension to $S-P$ interference has been
  studied for $B\to K\pi\ell^+\ell^-$
  \cite{Blake:2012mb,Becirevic:2011bp}, and for $\bar{B}\to
  \pi\pi\ell^-\bar\nu_\ell$ \cite{Faller:2013dwa}. For a first study
  of $S$, $P$
  and $D$ interference, see \cite{Das:2014sra}}.\\

@@

We now study the performance of the proposed method. In order to do so, we simulate
individual events for two separate physical processes: one uni-angular, and one tri-angular decay
distribution. We repeats theses analysis for varying sample sizes, ranging from
$50$ to $500$ events. Our toy analysis base on SM predictions for angular observables
in the decays $B\to K\ell^+\ell^-$ and $B\to K^*(\to K\pi)\ell^+\ell^-$, respectively.
In order to faithfully investigate the performance of the method of moments, we repeats our
numerical studies for several bin sizes in the kinematic range $1 \GeV^2 \leq q^2 \leq 6\GeV^2$,
as well as $15 \GeV^2 \leq q^2 \leq q^2_\text{max}$. Here, the bin size is chosen either
as $1\GeV^2$ or $0.5\GeV^2$. This setup is meant to ensure that a wide spectrum of possible
values for the angular observables is investigated.\\

>>>>

\begin{abstract}
  \fred{Our goal is to obtain a complete set} of angular observables
  arising in a generic multi-body process without the need to carry
  out a likelihood fit of the angular distribution to the measured
  events. Instead, we \fred{apply the method of moments} that relies
  both on the orthogonality of angular functions and the estimation of
  integrals by Monte Carlo techniques. \fred{The main advantage of
    this method is that the joint distribution of all observables
    given the data can be easily extracted, even for very few
    events. Our main contribution is to show explicitly how systematic
    uncertainties from detector-resolution and acceptance effects can
    be incorporated in a general algorithm. Finally, we present the
    necessary process-dependent formulae needed for direct
    implementation of the method for several rare decays of
    interest.}
\end{abstract}

@@

Our initial motivation for studying what we wish to call the
\emph{method of moments} is the determination of angular observables
in the rare decay FCNC-mediated decay $\bar{B}\to \bar{K}^*(\to
\bar{K}\pi)\ell^+\ell^-$. However, note that the method we
describe in the following is general, and it applies to arbitrary decay
or scattering processes. We find a previous work \cite{Dighe:1998vk}
that advocates this method chiefly for the determination of angular
observables in non-leptonic $B$ decays, but also mentions the
applicability to semileptonic decays. Our aim is to improve upon this previous
work by studying the uncertainties that are introduced by mismodelling
of the angular distribution, and by working out a recipe to determine
and unfold detector effects. Moreover, we show
that the method of moments has two major advantages over the usual
approach based on likelihood fits:
\begin{enumerate}
    \item Likelihood fits have convergence problems for a small number of
        events, and can require reparametrizations and/or approximations
        for a successful fit to the signal PDF, see e.g. \cite{Egede:2008uy}.
        The method of moments has no such convergence problems.
    \item The results of likelihood fits suffer from systematic uncertainties
        and systematic bias (see e.g. \cite{Egede:2008uy}), especially in but not limited
        to cases in which the PDFs do not faithfully model the underlying physics. We will
        show that the method of moments does not suffer a very particular type of
        mismodelling that arises from introducing a cutoff in a partial wave expansion.
    \item Using the method of moments, the joint probability distribution of the
        angular observables converges towards a multivariate Gaussian distribution.
        This allows an easy transfer of correlation information from the experiments to
        interested theorists.
\end{enumerate}

We continue with basic definitions that pertain to angular observables, and our
results in the subsequent sections. Let $\thvec$ denote the set of all
angles, and let $\nuvec$ denote the set of all
other non-angular kinematic variables needed to fully specify the
final state of the process under study. For example, $\nuvec$ may
include invariant masses or center-of-mass energies. We define an
angular observable $S_i$ as a coefficient in the probability density
function (PDF) of the process by means of
\begin{align}
    \label{eq:def-P}
    P(\nuvec, \thvec) \equiv \sum_i S_i(\nuvec) \times f_i(\thvec)\,.
\end{align}
Here, the dependence on the decay angles $\thvec$ has been
explicitly factored out in terms of the angular functions
$\{f_i(\thvec)\}$. We assume there exists a dual basis of functions
$\{\dual{f}_i(\thvec)\}$ such that the orthonormality relations
\begin{equation}
    \label{eq:def-ortho-rel}
    \int_\Omega \rmdx{\vec{\theta}} \dual{f}_i(\thvec) f_j(\thvec)  = \delta_{ij}
\end{equation}
hold with $\Omega$ representing the full angular phase space relevant
to the process.  For particle decays, $P$ is generally expressed in
terms of the fully differential decay width,
\begin{align}
    \label{eq:def-P-decay}
    P(\nuvec, \thvec) \equiv \frac{1}{\Gamma}\frac{\rmdx{^{2}\Gamma}}{\rmdx{\vec\nu} \rmdx{\vec\theta}}\,,
\end{align}
where $\Gamma$ is the total decay width. For a scattering process, one can similarly use
\begin{align}
    \label{eq:def-P-scattering}
    P(\nuvec, \thvec) \equiv \frac{1}{\sigma}\frac{\rmdx{^{2}\sigma}}{\rmdx{\vec\nu} \rmdx{\vec\theta}}\,,
\end{align}
where the total cross section $\sigma$ is used for the
normalization. Since the determination of the total decay width or
total cross section can be quite difficult, we emphasize that
different normalizations for $P$ can be used.  For instance, the total
decay width (or cross section) of the process of interest can be
replaced by the corresponding quantity of a control-channel
process. This change of normalization is equivalent to a linear rescaling of
the angular observables $\lbrace S_i\rbrace$; thus ratios or similar suitable combinations
of the angular observables are not affected by this change of normalization.\\

Our method is an extension of the classical method of moments with
orthogonal functions~\cite[sec. 8.2]{James:2006zz}. The only difference
is that conventionally the angular functions are assumed
\emph{self-dual}, $\dual{f}_i = f_i$. However, it suffices that the
system of angular functions $\{f_i(\thvec)\}$ \emph{can} be transformed into
an orthonormal basis. It turns out to be much more convenient to work
in the basis of Legendre polynomials but the statistical properties
are equivalent for both approaches provided that one replace $f_i \to
\dual{f}_i$ appropriately. Using the ansatz
\begin{equation}
  \label{eq:dual-ansatz}
  \dual{f}_i = \sum_{j} a_{ij} f_j \,,
\end{equation}
the dual basis needs to be worked out case by case through solving the
linear system of equations (\ref{eq:def-ortho-rel}). For a selection
of hadron decays with a $b$ quark in the initial state and two leptons
in the final state, we list the dual bases in a series of appendices
\ref{app:btokll} through \ref{app:lambdabtolambdall}. Note that a
similar analysis was done in~\cite{Dighe:1998vk} for the decays $B \to
J/\psi \phi$ and $B \to J/\psi K^{*}$.

@@

The orthonormality relations \refeq{def-ortho-rel} imply that a single angular observable $S_i$
can be projected out of the full PDF $P$ by means of
\begin{equation}
    \label{eq:det-Pi-analytical}
    S_i(\nuvec) = \int_{\Omega} \rmdx{\vec\theta}  P(\nuvec, \thvec) \dual{f}_i(\thvec)\,.
\end{equation}
where $\lbrace \dual{f}_i \rbrace$ denotes a \emph{dual basis} of angular functions. In general, $\lbrace \dual{f}_i \rbrace$ may differ from $\lbrace f_i \rbrace$. This is the
case for our selection of applications in appendices \ref{app:btokll} through \ref{app:btokstarll}.\\

@@

Integration over the non-angular variables yields
\begin{equation}
    \langle S_i\rangle
    \equiv \int \rmdx{\vec\nu} S_i(\nuvec)
    = \int \rmdx{\vec \nu} \left[\int_{\Omega} \rmdx{\vec \theta} P(\nuvec,\thvec) \dual{f}_i(\thvec) \right].
\end{equation}
The remainder of this section describes the method of moments, in
which we replace the analytical integration by Monte Carlo (MC)
estimates.  The central tenet of MC integration is the fact that the
expectation value $E_P[g]$ of some function $g(x)$ under the
probability density $P(x)$,
\begin{equation}
    E_P[g] \equiv \int \rmdx{x} P(x) g(x),
\end{equation}
can be approximated by the consistent and unbiased
estimator $\est{E_P[g]}$~\cite[sec. 8.2]{James:2006zz}
\begin{equation}
    \label{eq:mc-id}
    E_P[g] \to \widehat{E_P[g]} \equiv \frac{1}{N} \sum_{n=1}^N g(x^{(n)}) \,,\,    x^{(n)} \sim P
\end{equation}
due to the strong law of large numbers for $N \to \infty$, assuming
that the variates $x^{(n)}$, $n = 1, \dots, N$, are distributed as
$P$.
Throughout this letter we denote all MC estimators with a wide hat.\\

Application of \refeq{mc-id} then yields
\begin{equation}
    \langle S_i\rangle \to \widehat{\langle S_i\rangle} = \frac{1}{N} \sum_{n=1}^{N} \dual{f}_i(x^{(n)})\,.
\end{equation}
It is often of interest to obtain observables integrated over certain
bins of $\nuvec$. We define
\begin{align}
    \langle S_i\rangle_{\vec{a},\vec{b}}
    & \equiv \int_{\vec{a}}^{\vec{b}} \rmdx{\nuvec} S_i(\nuvec)\\
    & = \int_{\vec{a}}^{\vec{b}} \rmdx{\nuvec} \left[\int_{\Omega} \rmdx{\vec\theta} P(\nuvec,\thvec) \dual{f}_i(\thvec)  \right]\\
    & = \int \rmdx{\nuvec} \left[\int_{\Omega} \rmdx{\vec\theta}  P(\nuvec,\thvec) \dual{f}_i(\thvec)
        \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,
        \right],
\end{align}
where the argument of the indicator function $ \mathbf{1}(\vec{a} \le
\nuvec \le \vec{b})$ is to be interpreted componentwise.
Application of \refeq{mc-id} immediately yields
\begin{equation}
    \label{eq:bin-importance}
    \widehat{\langle S_i\rangle_{\vec{a},\vec{b}}}
    = \frac{1}{N} \sum_{n=1}^{N} \dual{f}_i(x^{(n)})         \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,.
\end{equation}
For notational simplicity, let us forget about the $\nuvec$
integration and consider only $S_i$. In the limit $N \to \infty$, the
central limit theorem (CLT) implies that the random vector
\begin{equation}
  \label{eq:angular-obs-vec}
  \vecest{S} \equiv (\est{S_0}, \dots, \est{S_i}\,,
  \dots)
\end{equation}
follows a multivariate Gaussian distribution $\mathcal{N}(\vec{S},
\widehat{\Sigma})$ centered on the true value $\vec{S}$ with the
covariance $\widehat{\Sigma}_{ij}$ estimated as
\begin{equation}
    \Sigma_{ij} \equiv \cov[S_i,S_j] \to \est{\Sigma_{ij}} \equiv \est{\cov}[{S}_i, {S}_j]
        = \frac{1}{N - 1} \sum_{n=1}^{N} \Big[\dual{f}_i\big(x^{(n)}\big) - \est{S_i}\Big]\,\Big[\dual{f}_j\big(x^{(n)}\big) - \widehat{S_j}\Big]\,.
\end{equation}
In our physics applications, the parameter space is compact and each
$\dual{f}_i$ is bounded. Hence the requisites for the most basic
version of the CLT to hold --- finite mean and covariance of
$\dual{f}_i$ --- are automatically satisfied.
In our numerical analysis the sample covariance rapidly converges towards the true covariance
matrix; see also \refsec{numerics}.\\

Compared to the usual maximum-likelihood approach, we find for the method of moments:
\begin{enumerate}
  \item The angular observable ${S_i}$ can be determined
  independently of any other observable ${S_j}$. It is therefore
  much more robust to physics assumptions needed to define the full
  likelihood. In particular, this means one does not have to be
  specific regarding the form of new-physics contributions; in fact,
  one does not even need to be able to explicitly formulate the
  likelihood at all.
\item It is superior for a small number of samples $N$. Likelihood
  fits tend to be numerically unstable if lots of parameters need to
  be estimated from sparse data. This is more severe if the mode of
  the likelihood is near the boundary of the physically allowed
  region.\fred{Is there some LHCb note we could cite?} 
  For some of these decays of interest, there are only $\order{10}$ of events
  per bin on record.
\item The estimate is unbiased for any $N$. In contrast, the
  maximum-likelihood estimate has a bias of order
  $1/N$~\cite{Cox:1968}. In practice, one should keep in mind the
  bias-variance trade-off: it is a well known phenomenon that removing
  the bias usually leads to an increase in variance of the sampling
  distribution of the estimator and vice versa~\cite[sec. 7.3]{James:2006zz}. From a
  Bayesian decision-theory point of view, both contribute similarly to
  the expected loss associated with deciding on just one value of the
  unknown parameter. It is therefore highly questionable to reduce
  only the bias~\cite[sections 13.8,17.2]{jaynes:2003}. In fact, for
  the results discussed below in \refsec{numerics}, the likelihood
  fits --- if they converge --- produce uncertainties (10--15)\%
  smaller than those from the method of moments.
\item The approximate multivariate Gaussian distribution of $\vecest{S}$
  allows easier and more correct transfer of the information in the
  data to interested theorists for more accurate fits of
  standard-model and new-physics parameters \cite{Altmannshofer:2013foa,
    Descotes-Genon:2013wba,Bobeth:2013soa},
  or for more precise predictions of optimized observables; see e.g.,
  \cite{Egede:2008uy,Egede:2010zc,Bobeth:2010wg,Becirevic:2011bp,
    Bobeth:2012vn,Matias:2012xw,DescotesGenon:2012zf} for definitions
  of such optimized observables in $B\to K^*\ell^+\ell^-$ decays,
  \cite{Faller:2013dwa} for application to the decay $B\to
  \pi\pi\ell^-\bar\nu_\ell$, and \cite{Boeer:2014xx} for observables
  in $\Lambda_b\to\Lambda(\to N\pi)\ell^+\ell^-$). While the
  likelihood also approaches a multivariate Gaussian as $N \to
  \infty$, the two methods differ in their utility as input for
  theorists if $\widehat{\vec{S}}$ is not well
  inside the physical region. For example, suppose there are two
  angular observables that are constrained to a triangular region by
  phase-space or unitarity arguments as
  \begin{equation}
    \label{eq:constr-ex}
    |S_1| + S_2 \le 1, \, S_1 \in [-1,1],\, S_2 \in [0,1] \,.
  \end{equation}
  It may (and often does) happen in practice that $\widehat{\vec{S}}$
  is close or even outside the allowed region such that a significant
  part of the probability mass covers unphysical values. In a Bayesian
  fit, one would take $\mathcal{N}(\vecest{S} | \vec{S},
  \widehat{\Sigma})$ as the sampling distribution of the ``data'', and
  simply set a uniform prior on the triangle in the $\vec{S}$ plane
  defined by \refeq{constr-ex} to have a well defined problem. This
  could be trivially combined with other independent information in a
  global fit. Someone with a different physics model might have to
  consider a
  different physical region, and could incorporate it just as easily.\\

  For a likelihood fit, the constraint needs to be part of the
  analysis performed by the experimental collaboration, and the
  resulting likelihood as a function of $\vec{S}$ is distinctly not
  Gaussian.  For experimentalists, communicating such a result has
  proved to be challenging due to technical reasons (data formats,
  size etc.) and the fear that theorists misinterpret the many
  nuisance parameters (from control channels, detector calibrations
  etc.) that are also part of the likelihood. This leads to the
  undesirable situation that only the mode and standard errors
  are reported, and theorists often include the results as independent
  measurements with a Gaussian distribution and disregard the boundary
  problem as well as correlations altogether.

@@

In several interesting processes we might only have an approximate
result for $P$.  In this section, we focus on one particular class of
mismodelling of the signal PDF: the angular-momentum cutoff in partial
wave expansions. This mismodelling potentially affects a large number
of decays and scattering processes. For the sake of clarity we take
the interesting class of four-body decays $B\to P_1 P_2 \ell_1$
as an example\footnote{This includes the rare $b\to s$ mediated $B$ decay $B
  \to K\pi\ell^+\ell^-$, and the $V_{ub}$ suppressed decay $B\to
  \pi\pi\ell^+\bar\nu_\ell$. For both examples the PDF $P$ is known in
  the small-width approximation and when assuming a pure $P$-wave,
  resonant final state. An extension to $S-P$ interference has been
  studied for $B\to K\pi\ell^+\ell^-$
  \cite{Blake:2012mb,Becirevic:2011bp}, and for $\bar{B}\to
  \pi\pi\ell^-\bar\nu_\ell$ \cite{Faller:2013dwa}. For a first study
  of $S$, $P$
  and $D$ interference, see \cite{Das:2014sra}
}.\\

@@

We now study the performance of the proposed method. In order to do so, we simulate
individual events for two separate physical processes: one uni-angular, and one tri-angular decay
distribution. We repeats theses analysis for varying sample sizes, ranging from
$50$ to $500$ events. Our toy analysis base on SM predictions for angular observables
in the decays $B\to K\ell^+\ell^-$ and $B\to K^*(\to K\pi)\ell^+\ell^-$, respectively.
In order to faithfully investigate the performance of the method of moments, we repeats our
numerical studies for several bin sizes in the kinematic range $1\,\GeV^2 \leq q^2 \leq 6\,\GeV^2$,
as well as $15\,\GeV^2 \leq q^2 \leq q^2_\text{max}$. Here, the bin size is chosen either
as $1\,\GeV^2$ or $0.5\,\GeV^2$. This setup is meant to ensure that a wide spectrum of possible
values for the angular observables is investigated.\\
